---
title: "Portfolio 5/SocKultExam"
author: "Esther and Simon"
date: "3 maj 2018"
output: html_document
---

Load data, check BOT/Partner behaviour, make preliminary plots

```{r}
# Working directory
setwd("C:/Users/simon/Google Drev/SocKult exam/Portfolio5/Data")

# Load libraries
library(dplyr);library(ggplot2);library(rethinking);library(brms);library(tibble);library(boot);library(cowplot)

# Load data
temp = list.files(pattern ="*.csv")
d=data.frame()
for (i in 1:length(temp)) {
  csv=read.csv(temp[i])
  csv$X = csv$X + 1
  d=rbind(d, csv)
}

# Rename column
colnames(d)[colnames(d) == "X"] <- "Trial"

# Recode PR
d$PR = as.character(d$PR)
d$PR[d$PR=="no"] = "0"
d$PR[d$PR=="yes"] = "1"
d$PR = as.numeric(d$PR)

# Check Bot behaviour
d_yes = d[d$PR == 1,]
d_yes$BR = as.character(d_yes$BR)
d_yes$BR[d_yes$BR=="Defect"] = "0"
d_yes$BR[d_yes$BR=="Cooperate"] = "1"
d_yes$BR = as.numeric(d_yes$BR)

# A = 27,5 % cooperation
BOTA=d_yes[d_yes$BOT == "A",]
sum(BOTA$BR)/length(BOTA$BR)

# B = 66 % cooperation
BOTB=d_yes[d_yes$BOT == "B",]
sum(BOTB$BR)/length(BOTB$BR)

# C = 27.6 % cooperation
BOTC=d_yes[d_yes$BOT == "C",]
sum(BOTC$BR)/length(BOTC$BR)

# D = 72 % cooperation
BOTD=d_yes[d_yes$BOT == "D",]
sum(BOTD$BR)/length(BOTD$BR)

# Duplicate Bot column 
d$Partner = d$BOT

# Overall cooperation rate for each partner
ggplot(d, aes(Partner, PR, fill = Partner)) + geom_bar(stat= "summary") + geom_errorbar(stat="summary", fun.data=mean_se, width=0.3) + theme(legend.position= "none") + ggtitle("Player cooperation for each partner") + ylab("Player cooperation")

# Player response as a function of time
ggplot(d, aes(x = Trial, y= PR, colour = Partner)) + geom_smooth(alpha = 0.2) + ggtitle("Player cooperation level as a function of time") + xlab("Time") + ylab("Player cooperation")
```


Primary models (Portfolio 5 and Social and Cultural dynamics)

```{r}
# Convert PR into a factor
d$PR=as.numeric(d$PR)

# Create 3 models predicting player response
m1 <- brm(PR ~ Trial*Social_Rep*Prog_Beh  + (1+Trial*Social_Rep*Prog_Beh|ID), 
            data = d, family = bernoulli(), cores = 2, chains = 4, iter = 1e3)

m2 <- brm(PR ~ Trial+Social_Rep+Prog_Beh  + (1+Trial+Social_Rep+Prog_Beh|ID), 
            data = d, family = bernoulli(), cores = 2, chains = 4, iter = 1e3)

m3 <- brm(PR ~ Trial*Prog_Beh  + (1+Trial*Prog_Beh|ID), 
            data = d, family = bernoulli(), cores = 2, chains = 4, iter = 1e3)

# Summary for each model
summary(m1)
summary(m2)
summary(m3)

# To add WAIC onto each model
m1<-add_ic(m1,ic="waic")
m2<-add_ic(m2,ic="waic")
m3<-add_ic(m3,ic="waic")

# Compare models 
compare_ic(m1,m2,m3, ic="waic")

# Compare estimates from each model
fixef(m1)
fixef(m2)
fixef(m3)

# Convert brms fit into matrix
fit1 = as.matrix(m1)
fit2 = as.matrix(m2)
fit3 = as.matrix(m3)

# Create visual overview of model parameters for each model
mcmc_intervals(fit1, pars = c("b_Intercept", "b_Trial", "b_Social_RepLow", "b_Prog_BehGood", "b_Trial:Social_RepLow", "b_Trial:Prog_BehGood", "b_Social_RepLow:Prog_BehGood", "b_Trial:Social_RepLow:Prog_BehGood"), 
           prob = 0.95) +ggplot2::ggtitle("Model 1 parameters")

mcmc_intervals(fit2, pars = c("b_Intercept", "b_Trial", "b_Social_RepLow", "b_Prog_BehGood"), 
           prob = 0.95) +ggplot2::ggtitle("Model 2 parameters")

mcmc_intervals(fit3, pars = c("b_Intercept", "b_Trial", "b_Prog_BehGood", "b_Trial:Prog_BehGood"), 
           prob = 0.95) +ggplot2::ggtitle("Model 3 parameters")

# To get posterior probability and plot marginal effects
pp_check(m1) 

d$PR=as.factor(d$PR)

plot(marginal_effects(m1)) 

# To get model predictions
predictions=predict(m1, type = "response")
d$prediction = predictions[,1]

# To make a plot comparing model predictions with actual data. Actual data = red, predictions = blue
dens(d$PR, col = "red") 
dens(d$predict, col = "blue", add = TRUE)

# To create a confusion matrix to get accuracy
d$Predictions[d$prediction>0.5]=1 
d$Predictions[d$prediction<=0.5]=0
confusionMatrix(data = d$Predictions, reference = d$PR)

# Plot predictions
pred_plot=ggplot(d, aes(y=Predictions, x= Trial, colour = Partner)) + geom_smooth() + ggtitle("Predictions of model 1") + theme(legend.position= "none") + xlab("Time") + ylim(0, 1)

# Plot data
data_plot=ggplot(d, aes(x = Trial, y= PR, colour = Partner)) + geom_smooth(alpha = 0.2) + ggtitle("Player cooperation level as a function of time") + xlab("Time") + ylab("Player cooperation") + theme(legend.position= "none") + ylim(0, 1)

# Combine the two plots
plot_grid(pred_plot, data_plot)

```


Further analysis for Social and Cultural dynamics exam: 
- Predicting indidual ratings from random effects

```{r}
# Extract random effect. Individual variability.
random=ranef(m1) 
random=as.data.frame(random)
random=tibble::rownames_to_column(random)

# Change column names
colnames(random)[colnames(random) == "rowname"] <- "ID"
colnames(random)[colnames(random) == "ID.Estimate.Intercept"] <- "Intercept_effect"
colnames(random)[colnames(random) == "ID.Estimate.Social_RepLow"] <- "SocialRep_Low"
colnames(random)[colnames(random) == "ID.Estimate.Prog_BehGood"] <- "ProgBeh_Good"
colnames(random)[colnames(random) == "ID.Estimate.Social_RepLow.Prog_BehGood"] <- "SocialRep_Low:ProgBeh_Good"
colnames(random)[colnames(random) == "ID.Estimate.Trial"] <- "Trial"
colnames(random)[colnames(random) == "ID.Estimate.Trial.Prog_BehGood"] <- "Trial:ProgBeh_Good"

# Create plots with individual variability most important effects
plot1=ggplot(random, aes(Intercept_effect)) + geom_density()
plot2=ggplot(random, aes(SocialRep_Low)) + geom_density()
plot3=ggplot(random, aes(ProgBeh_Good)) + geom_density()
plot4=ggplot(random, aes(SocialRep_Low:ProgBeh_Good)) + geom_density()
plot5=ggplot(random, aes(Trial)) + geom_density()
plot6=ggplot(random, aes(Trial:ProgBeh_Good)) + geom_density()

# Combine plots
cowplot::plot_grid(plot1, plot2, plot3, plot4, plot5, plot6)

# A lot of individual variability. Could be due to lack of data, but is also likely to be the results of individual differences. 

# To get rating for each participant
d_rating=aggregate(d[, 12:15], list(d$ID), mean)

# Change column name
colnames(d_rating)[colnames(d_rating) == "Group.1"] <- "ID"

# Merge dataset
d_rating2 = merge(d_rating, random, by = "ID")

# Predicting rating from random effects
m_rating <- brm(cbind(A_rating, B_rating, C_rating, D_rating) ~ Intercept_effect + SocialRep_Low + ProgBeh_Good+SocialRep_Low:ProgBeh_Good, 
            data = d_rating2, family = gaussian(), cores = 2, chains = 4, iter = 1e3)

# Summary and plot of rating model
summary(m_rating)
plot(m_rating)

# Make rating based on model and convert to data frame
pred_rating=predict(m_rating)
pred_rating=as.data.frame(pred_rating)

# Save to ratings datafile 
d_rating2$PredA_rating = pred_rating$Estimate.Arating
d_rating2$PredB_rating = pred_rating$Estimate.Brating
d_rating2$PredC_rating = pred_rating$Estimate.Crating
d_rating2$PredD_rating = pred_rating$Estimate.Drating

#Calculating rmse
rmse(d_rating2$A_rating, d_rating2$PredA_rating)
rmse(d_rating2$B_rating, d_rating2$PredB_rating)
rmse(d_rating2$C_rating, d_rating2$PredC_rating)
rmse(d_rating2$D_rating, d_rating2$PredD_rating)

# Plot of rating given by the players
Aplot=ggplot(d_rating2, aes(A_rating)) + geom_density()
Bplot=ggplot(d_rating2, aes(B_rating)) + geom_density()
Cplot=ggplot(d_rating2, aes(C_rating)) + geom_density()
Dplot=ggplot(d_rating2, aes(D_rating)) + geom_density()

plot_grid(Aplot, Bplot, Cplot, Dplot) # Combine plots

# Plot of predictions given by the model
PAplot=ggplot(d_rating2, aes(PredA_rating)) + geom_density()
PBplot=ggplot(d_rating2, aes(PredB_rating)) + geom_density()
PCplot=ggplot(d_rating2, aes(PredC_rating)) + geom_density()
PDplot=ggplot(d_rating2, aes(PredD_rating)) + geom_density()

plot_grid(PAplot, PBplot, PCplot, PDplot) #Combine plots
```


Further analysis for Social and Cultural dynamics exam: 
- Modelling reaction time for each partner and reaction time as a function of time

```{r}

# Make model with player cooperation for each partner
m4 <- brm(PR ~ 0 + Social_Rep:Prog_Beh  + (0+Social_Rep:Prog_Beh|ID), 
            data = d, family = bernoulli(), cores = 2, chains = 4, iter = 1e3)

# Make model with player reaction time for each partner
m_RT <- brm(RT ~ 0 + Social_Rep:Prog_Beh  + (0+Social_Rep:Prog_Beh|ID), 
            data = d, family = exponential(), cores = 2, chains = 4, iter = 1e3)

summary(m4) # Same pattern as what is seen in the main models
summary(m_RT) # No effect of condition on reaction time

# To see the effect of trial on reaction time
ggplot(d, aes(x=Trial, y=RT, colour = BOT)) + geom_smooth()
ggplot(d, aes(x=Trial, y=RT)) + geom_smooth() + ggtitle("Reaction time (RT) as a function of Trial")

```

Further analysis for Social and Cultural dynamics exam: 
- Are we making decisions in an optimal bayesian manner?

```{r}

# STAN SETTINGS

N_CORES = 2
ITER = 1e3

STAN_CONTROL = list(
    adapt_delta = .8,
    max_treedepth = 10
)

# make 1 variable with logit of behaviour (.7) score
# make 1 variable with logit of reliability (divided by 5 before logit) score
d_bayes = d
d_bayes$l_behavior[d_bayes$Prog_Beh == "Good"] = logit(0.7)
d_bayes$l_behavior[d_bayes$Prog_Beh == "Bad"] = logit(0.3)  
d_bayes$l_reliability[d_bayes$BOT == "A"] = logit(3.7/5)
d_bayes$l_reliability[d_bayes$BOT == "B"] = logit(1.5/5)
d_bayes$l_reliability[d_bayes$BOT == "C"] = logit(1.7/5)
d_bayes$l_reliability[d_bayes$BOT == "D"] = logit(3.8/5)

# Creating a non weighted bayesian model
naive_f = bf(PR ~ bias + l_reliability + l_behavior,
             bias ~ 1 + (1|ID),
             nl = TRUE)

naive_prior = c(prior(normal(0,1), nlpar = "bias")) # Prior for bias)

naive_m = brm(naive_f,
              d_bayes,
              prior = naive_prior,
              family = "bernoulli",
              chains = N_CORES, cores = N_CORES, iter = ITER,
              control = STAN_CONTROL)


# Creating a weighted bayesian model


# define weighting function for use by stan
# here, a uses a log link and w uses a scaled logit link.
# ie, the raw a and w parameters in the model are unbounded in order for us to put predictors on them
# then the F function transforms them into (0, Inf) for a and (.5, 1) for w

F_stancode = "
real F3(real a_raw, real L_raw, real w_raw) {
real a;
real L;
real w;
a = exp(a_raw);
L = exp(L_raw * a);
w = 0.5 + inv_logit(w_raw)/2;
return log((w * L + 1 - w)./((1 - w) * L + w));
}
"

F3 <- function(a_raw, L_raw, w_raw) {
    # used by brms to back-translate
    a = exp(a_raw)
    L = exp(L_raw * a)
    w = .5 + inv.logit(w_raw) / 2
    log((w * L + 1 - w)/((1 - w) * L + w))
    
}

weighted_f = bf(PR ~ bias + F3(0, l_behavior, wBehavior) + F3(0, l_reliability, wReputation),
               wBehavior + wReputation + bias ~ 1 + (1|ID),
               nl = TRUE)

weighted_prior = c(prior(normal(0,1), nlpar = "wReputation"),
                   prior(normal(0,1), nlpar = "wBehavior"),
                   prior(normal(0,1), nlpar = "bias"))

weighted_m = brm(weighted_f,
              d_bayes,
              stan_funs = F_stancode,
              prior = weighted_prior,
              family = "bernoulli",
              chains = N_CORES, cores = N_CORES, iter = ITER,
              control = STAN_CONTROL)

# Adding WAIC for model comparison 
naive_m<-add_ic(naive_m,ic="waic")
weighted_m<-add_ic(weighted_m,ic="waic")

# Comparing models 
compare_ic(m1, naive_m, weighted_m, ic="waic")

```
